---
title: "Minnesota schools Open enrollment by race"
author: "MaryJo Webster"
date:  "Last updated: `r Sys.Date()`"
output:
  html_document: 
    #toc: true
    #to_depth: 1
    #toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#install.packages("rmarkdown")
#install.packages("RMySQL")

#setwd("C:/Users/webstmj/OneDrive - Star Tribune/Rprojects/schools_openenroll")


library(readr) #importing csv files
library(dplyr) #general analysis 
library(ggplot2) #making charts
library(lubridate) #date functions
library(reshape2) #use this for melt function to create one record for each team
library(tidyr)
library(janitor) #use this for doing crosstabs
library(scales) #needed for stacked bar chart axis labels
library(knitr) #needed for making tables in markdown page
library(htmltools)#this is needed for Rstudio to display kable and other html code
library(rmarkdown)
library(kableExtra)
library(ggthemes)
library(RMySQL)



#http://strib-data-public.s3-us-west-1.amazonaws.com/projects/schools/openenroll.html

```




```{r, code=readLines("openenroll_processdata.R"), echo=FALSE, warning=FALSE, message=FALSE}

```

# Total leaving home districts by year
```{r}
annual_totals <-  openenroll_bydistrict_final %>% group_by(yr) %>% summarise(leaveTrad=sum(LeavingToTrad),leaveCharter=sum(LeavingToCharter))

annual_totals <-  melt(annual_totals, id.vars="yr")

annual_totals$variable <-  factor(annual_totals$variable, levels=c("leaveTrad", "leaveCharter"), labels=c("Open Enroll", "Charter"))

fill <-  c("#5F9EA0", "#E1B378")

g1 <-  ggplot() + 
  geom_bar(aes(y=value, x=yr, fill=variable), data=annual_totals, stat="identity") + 
  theme(legend.position="bottom", legend.direction="horizontal", legend.title=element_blank())+
  scale_x_continuous(name="School year", breaks=seq(2000,2018,2))+
  scale_y_continuous(name="Number of students", limits=c(0,150000))+
  scale_fill_manual(values=fill)+
  theme_hc()+
      labs(title = "Number of students leaving home districts", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")

plot(g1)
```


#Percent of students leaving statewide
```{r}


annual_pct <- openenroll_bydistrict_final %>% group_by(yr) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g2 <- ggplot(data=annual_pct, aes(x = annual_pct$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .20), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
  theme_hc()+
        labs(title = "Pct of students leaving home districts", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")


plot(g2)
```


#Racial groups
```{r}
annual_pct_groups <- openenroll_bydistrict_final %>% group_by(yr, studentgroup) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g3 <- ggplot(data=annual_pct_groups, aes(x = annual_pct_groups$yr, y = pct, group=studentgroup, col=studentgroup)) +
                  geom_line(stat="identity",  size=1.5)+
  scale_y_continuous(name="Percent of residents leaving", limits=c(0, .3), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Pct of students leaving home districts", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g3)
```





#By location
```{r}
annual_pct_location <- openenroll_bydistrict_final %>% group_by(yr, Location) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g4 <- ggplot(data=annual_pct_location, aes(x = annual_pct_location$yr, y = pct, group=Location, col=Location)) +
                  geom_line(stat="identity",  size=1.5)+
  scale_y_continuous(name="Percent of residents leaving", limits=c(0, .35), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Pct of students leaving home districts", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g4)
```


#Core Cities only - by racial groups
```{r}
annual_pct_corecities <- openenroll_bydistrict_final %>% filter(Location=='CORE CITIES') %>% group_by(yr, studentgroup) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g6 <- ggplot(data=annual_pct_corecities, aes(x = annual_pct_corecities$yr, y = pct, group=studentgroup, col=studentgroup)) +
                  geom_line(stat="identity",  size=1.5)+
  scale_y_continuous(name="Percent of residents leaving", limits=c(0, .40), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Pct of students leaving home districts", 
       subtitle = "Core cities - 1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g6)
```




#Metro districts losing largest percentage in 2017-18
```{r}
openenroll_bydistrict_final %>% filter(yr=='2018', Metro7County=='yes') %>% group_by(Organization) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez) %>% arrange(desc(pct)) %>% select(Organization, pct) %>% 
  filter(pct>=.2) %>% 
    kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Metro districts losing largest percentage in 2016-17
```{r}
openenroll_bydistrict_final %>% filter(yr=='2017', Metro7County=='yes') %>% group_by(Organization) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez) %>% arrange(desc(pct)) %>% select(Organization, pct) %>% 
    filter(pct>=.15) %>% 
    kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


#Metro Districts with largest share coming in- 2017-18
```{r}
openenroll_bydistrict_final %>% filter(yr=='2018', Metro7County=='yes') %>% group_by(Organization) %>% summarise(In=sum(ComingIn), enroll=sum(enrolled)) %>% mutate(pct=In/enroll) %>% arrange(desc(pct)) %>% select(Organization, pct) %>% 
  filter(pct>=.2) %>% 
    kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


#Minnetonka
```{r}
minnetonka <-  openenroll_bydistrict_final %>% filter(Organization=='Minnetonka Public School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=In/enroll)

g5 <- ggplot(data=minnetonka, aes(x = minnetonka$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="red", size=1.5)+
  scale_y_continuous(name="Percent of enrolled from elsewhere", limits=c(0, .35), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent coming into Minnetonka", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g5)
```


#Mahtomedi
```{r}
mahtomedi <-  openenroll_bydistrict_final %>% filter(Organization=='Mahtomedi Public School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=In/enroll)

g7 <- ggplot(data=mahtomedi, aes(x = mahtomedi$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="red", size=1.5)+
  scale_y_continuous(name="Percent of enrolled from elsewhere", limits=c(0, .35), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent coming into Mahtomedi", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g7)
```

#Brooklyn Center - coming in

```{r}
brooklyncenter_in <-  openenroll_bydistrict_final %>% filter(Organization=='Brooklyn Center School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=In/enroll)

g10 <- ggplot(data=brooklyncenter_in, aes(x = brooklyncenter_in$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="red", size=1.5)+
  scale_y_continuous(name="Percent of enrolled from elsewhere", limits=c(0, .5), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent coming into Brooklyn Center", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g10)
```


#Brooklyn Center - leaving


```{r}
brooklyncenter <-  openenroll_bydistrict_final %>% filter(Organization=='Brooklyn Center School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=leaving/rez)

g9 <- ggplot(data=brooklyncenter, aes(x = brooklyncenter$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .4), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent leaving Brooklyn Center", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g9)
```



#Columbia Heights
```{r}
columbiaheights <-  openenroll_bydistrict_final %>% filter(Organization=='Columbia Heights Public School Dist') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=leaving/rez)

g8 <- ggplot(data=columbiaheights, aes(x = columbiaheights$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .4), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent leaving Columbia Heights", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g8)
```



#Brooklyn Center -- top destinations
```{r}
openenroll %>% filter(residentdistrictnumber=='0286', datayear=='17-18') %>% group_by(districtname) %>% summarise(leavingto=sum(CountOfStudentsEnrolled)) %>% arrange(desc(leavingto)) %>% 
  filter(leavingto>=20) %>% 
    kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Brooklyn Center -- top origins

```{r}
openenroll %>% filter(districtnumber=='0286', datayear=='17-18') %>% group_by(residentdistrictname) %>% summarise(In=sum(CountOfStudentsEnrolled)) %>% arrange(desc(In)) %>% 
  filter(In>=10) %>% 
    kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


#Osseo
```{r}
osseo <-  openenroll_bydistrict_final %>% filter(Organization=='Osseo Public School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=leaving/rez)

g11 <- ggplot(data=osseo, aes(x = osseo$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .4), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent leaving Osseo", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g11)
```

#Minneapolis

```{r}
mpls <-  openenroll_bydistrict_final %>% filter(Organization=='Minneapolis Public School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=leaving/rez)

g12 <- ggplot(data=mpls, aes(x = mpls$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .4), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent leaving Minneapolis", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g12)
```


```{r}
annual_pct_groups_mpls <- openenroll_bydistrict_final %>% filter(Organization=='Minneapolis Public School District') %>% group_by(yr, studentgroup) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g14 <- ggplot(data=annual_pct_groups_mpls, aes(x = annual_pct_groups_mpls$yr, y = pct, group=studentgroup, col=studentgroup)) +
                  geom_line(stat="identity",  size=1.5)+
  scale_y_continuous(name="Percent of residents leaving", limits=c(0, .5), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Racial groups leaving Minneapolis", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g14)

#write.csv(annual_pct_groups_mpls, "mpls_by_race.csv", row.names = FALSE)
```




# St Paul
```{r}
sp <-  openenroll_bydistrict_final %>% filter(Organization=='St. Paul Public School District') %>% group_by(yr) %>%  summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), In=sum(ComingIn), enroll=sum(enrolled), rez=sum(residents)) %>% mutate(pct=leaving/rez)

g13 <- ggplot(data=sp, aes(x = sp$yr, y = pct, group=1)) +
                  geom_line(stat="identity", color="blue", size=1.5)+
  scale_y_continuous(name="Percent", limits=c(0, .4), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Percent leaving St. Paul", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g13)
```


```{r}
annual_pct_groups_sp <- openenroll_bydistrict_final %>% filter(Organization=='St. Paul Public School District') %>% group_by(yr, studentgroup) %>% summarise(leaving=sum(LeavingToTrad)+sum(LeavingToCharter), rez=sum(residents)) %>% mutate(pct=leaving/rez)



g15 <- ggplot(data=annual_pct_groups_sp, aes(x = annual_pct_groups_sp$yr, y = pct, group=studentgroup, col=studentgroup)) +
                  geom_line(stat="identity",  size=1.5)+
  scale_y_continuous(name="Percent of residents leaving", limits=c(0, .5), labels=percent)+
  scale_x_continuous(name="School year", breaks=seq(2000, 2018, 2))+
    theme_hc()+
      labs(title = "Racial groups leaving St. Paul", 
       subtitle = "1999-00 to 2017-18",
       caption = "Graphic by MaryJo Webster")
plot(g15)
```



```{r loadtoaws, echo=FALSE, eval=FALSE, results="hide"}

library(aws.s3)
Sys.setenv("AWS_ACCESS_KEY_ID" =  rstudioapi::askForPassword("AWS_ACCESS_KEY_ID"),

           "AWS_SECRET_ACCESS_KEY" = rstudioapi::askForPassword("AWS_SECRET_ACCESS_KEY"))

get_bucket("strib-data-public")


put_object(file = "openenroll.html", object = "projects/schools/openenroll.html", bucket = "strib-data-public")

put_object(file = "openenroll.html", object = "projects/schools/openenroll.html", bucket = "strib-data-public", acl=c("public-read"))

```

